Illustration of Adaptive Shrinkage

### 2016-09-30 analysis/adaptive_shrinkage.Rmd
Updated code in adaptive_shrinkage.Rmd up to plot_shrunk_vs_obs.

We simulate from two scenarios: in the first scenario, the effects are more
peaked about zero (**sim.spiky**); in the second scenario, the effects are
less peaked (**sim.bignormal**).

@@

sim.bignormal =
  rnormmix_datamaker(args = list(g         = normalmix(1,0,4),
                                 min_pi0   = 0,
                                 max_pi0   = 0,
                                 nsamp     = NSAMP,
                                 betahatsd = s))
```

@@

Here we plot the shrunken estimates against the observed values, colored
according to the (square root of) precision: precise esitmates being colored
red, and less precise estimates being blue. Two key features of the plots
illustrate the ideas of adaptive shrinkage: i) the estimates under the spiky
scenario are shrunk more strongly, illustrating that shrinkage adapts to the underlying distribution of beta; ii) in both cases, estimates with large
standard error (blue) are shrunk more than estimates with small standard
error (red) illustrating that shrinkage adapts to measurement precision.

```{r plotting}
make_df_for_ashplot = function(sim1,sim2,ash1,ash2,name1="spiky",name2="big-normal"){
  df = data.frame(betahat = c(sim1$input$betahat,sim2$input$betahat),
                beta_est = c(get_pm(ash1), get_pm(ash2)),
                lfsr = c(get_lfsr(ash1), get_lfsr(ash2)),
                s=c(sim1$input$sebetahat,sim2$input$sebetahat),
                scenario=c(rep(name1,NSAMP),rep(name2,NSAMP)))
  df$signif = df$lfsr<0.05
  return(df)
}

ashplot=function(df,xlab="Observed beta-hat",ylab="Shrunken beta estimate"){
  ggplot(df,aes(x=betahat,y=beta_est,color=1/s)) +
    xlab(xlab) +
    ylab(ylab) + geom_point() +  facet_grid(.~scenario) +   geom_abline(intercept=0,slope=1) +
  scale_colour_gradient2(midpoint = median(1/s),low="blue", mid="white",high="red",space="Lab")  +coord_fixed(ratio=1)
}
df=make_df_for_ashplot(sim.spiky,sim.bignormal,beta.spiky.ash,beta.bignormal.ash)
ashplot(df)
```

Now plot lfsr against z scores, colored
according to the (square root of) precision

```{r plotting_lfsr}

@@

z_lfsr_plot(df)
```

>>>>

We simulate from two scenarios: in the first scenario, the effects are more
peaked about zero (**sim.spiky**); in the second scenario, the effects are
less peaked at zero (**sim.bignormal**). A summary of the two data sets is
printed at the end of this chunk.

@@

sim.bignormal =
  rnormmix_datamaker(args = list(g         = normalmix(1,0,4),
                                 min_pi0   = 0,
                                 max_pi0   = 0,
                                 nsamp     = NSAMP,
                                 betahatsd = s))

cat("Summary of observed beta-hats:\n")
print(rbind(spiky     = quantile(sim.spiky$input$betahat,seq(0,1,0.1)),
            bignormal = quantile(sim.bignormal$input$betahat,seq(0,1,0.1))),
      digits = 3)

```

@@

Next we plot the shrunken estimates against the observed values, colored
according to the (square root of) precision: precise estimates being colored
red, and less precise estimates being blue. Two key features of the plots
illustrate the ideas of adaptive shrinkage: i) the estimates under the spiky
scenario are shrunk more strongly, illustrating that shrinkage adapts to the underlying distribution of beta; ii) in both cases, estimates with large
standard error (blue) are shrunk more than estimates with small standard
error (red) illustrating that shrinkage adapts to measurement precision.

```{r plot_shrunk_vs_obs}
make_df_for_ashplot =
  function (sim1, sim2, ash1, ash2, name1 = "spiky", name2 = "big-normal") {
    n    = length(sim1$input$betahat)
    lfsr = c(get_lfsr(ash1),get_lfsr(ash2))
    return(data.frame(betahat  = c(sim1$input$betahat,sim2$input$betahat),
                      beta_est = c(get_pm(ash1), get_pm(ash2)),
                      lfsr     = lfsr,
                      s        = c(sim1$input$sebetahat,sim2$input$sebetahat),
                      scenario = c(rep(name1,n),rep(name2,n)),
                      signif   = lfsr < 0.05))
  }

ashplot = function(df,xlab="Observed beta-hat",ylab="Shrunken beta estimate") {
  ggplot(df,aes(x = betahat,y = beta_est,color = 1/s)) +
    xlab(xlab) + ylab(ylab) + geom_point() + facet_grid(.~scenario) +   geom_abline(intercept = 0,slope = 1,linetype = "dotted") +
    scale_colour_gradient2(midpoint = median(1/s),low = "blue",mid = "white",
                           high = "red",space = "Lab") +
    coord_fixed(ratio = 1)
 }

df = make_df_for_ashplot(sim.spiky,sim.bignormal,beta.spiky.ash,
                         beta.bignormal.ash)
print(ashplot(df))
```

Now plot lfsr against z scores, colored according to the (square root of)
precision.

```{r plot_lfsr}

@@

print(z_lfsr_plot(df))
```
